Leaf mass per area and plant size drive differences in above-ground biomass distribution among woody plant functional types

### 2016-02-25 manuscript.Rnw
Clarify Eq. 1 and 2

% This next refers to stem xs-area, but then the equation is for diameter. Did we really use both?
In our dataset, stem cross-sectional area was measured either at breast height (1.3 - 1.34m) and/or at the base of the plant. In our analyses, we used basal stem area because many plants (N=\Sexpr{nrow(dataset[dataset$h.t < 1.3,])}) were \textless{} 1.3m tall. We therefore converted records where area was measured at breast height only to area at base, using a relationship fitted to individuals where both measurements existed.  For the subset of the data where both were measured, we estimated basal stem area ($\as$) from breast height stem area ($\asb$) from the equation,
\begin{equation}
D_{\rm S} = D_{\rm Sbh} \left(\frac{H}{H - H_{\rm bh}}\right)^c
\end{equation}
where $D_{\rm S}$ is the basal stem diameter (m), $D_{\rm Sbh}$ stem diameter at breast height, $H_{bh}$ the height at which $D_{\rm Sbh}$ was measured (typically 1.3 or 1.34m). The parameter $c$ was further expressed as a function of plant height:
\begin{equation}
c = c_0H^{c_1}.
\end{equation}
The estimated coefficients were $c_0$ = \Sexpr{signif(coef(basalafit)[1],3)}, $c_1$ = \Sexpr{signif(coef(basalafit)[2],3)}, root-mean square error = \Sexpr{signif(summary(basalafit)$sigma,3)}. The relationship was fit using a total of \Sexpr{length(residuals(basalafit))} observations covering the three major PFTs.

@@

From the variance partitioning (Table \ref{table:varparttable}), it is evident that $\afas$ did not vary appreciably between PFTs. Indeed, when all PFTs were plotted together, similar $\af$ was attained at a given $\as$ across the entire plant size range (with the exception of extremely large trees with \textgreater{} 1m$^2$ stem basal area, for which comparatively little data was available) (Fig. 4). We fit a GAM to $\af$ as a function of $\as$ with and without PFT as a covariate
% please clarify next line
(both as intercept and slope interaction),
which showed PFT significantly improved model fit (likelihood ratio test, p \textless{} 0.001). Nevertheless, the $R^2$ of the relationship only increased from \Sexpr{round(100*afas$R2a,2)} to \Sexpr{round(100*afas$R2b,2)}\%, indicating that while significant, the PFT had little additional predictive power.

@@

% This paragraph has been substantially changed, please review
We have confirmed the hypothesis that the amount of leaf area supported per plant is less variable between PFTs than the amount of leaf mass. \citet{poorter2015} hypothesised such a result might arise, because forests would converge to a similar leaf area index (LAI). Although we found remarkably similar leaf area per unit above-ground woody biomass ($\mfms$) between the three PFTs studied (Fig. S2), this does not in itself imply that LAI is equally similar. Indeed, in a global compilation of LAI for undisturbed forest stands, \citet{iio2014} reported a much higher LAI for evergreen gymnosperms compared to evergreen angiosperms, with deciduous angiosperms having intermediate and more variable LAI.  Moreover, we would in fact expect LAI to change with LMA, not because of changes in the amount of leaf area maintained by individual plants, but because LMA influences the shade-tolerance of seedlings \citep{poorter2006, falster2011}, and thereby the LAI of the community \citep{falster2011}. The ultimate reason LMA might effect shade tolerance is because is via its correlation with leaf lifespan \citep{wright2004}. For whole stands, we can therefore expect total stand leaf mass to be correlated with LMA, both because of of greater leaf mass per plant \citep[e.g.]{reich1992}, and the greater density of plants arising from increased leaf lifespan. Previous interpretations of differences in $\mfms$ between gymnosperms and angiosperms have focussed on differences in leaf life-span (LL) \citep{enquistniklas2002, poorter2012}. We argue that there are in fact two effects occurring simultaneously: differences in LMA cause differences in the $\mfms$ of individual plants, while differences in LL (and thereby LMA) cause differences in the LAI of whole stands.

>>>>

In our dataset, stem cross-sectional area was measured either at breast height (1.3 - 1.34m) and/or at the base of the plant. In our analyses, we used basal stem area because many plants (N=\Sexpr{nrow(dataset[dataset$h.t < 1.3,])}) were \textless{} 1.3m tall. We therefore converted records where area was measured at breast height only to area at base, using a relationship fitted to individuals where both measurements existed.  For the subset of the data where both were measured, we estimated basal stem area ($\as$) from breast height stem area ($\asb$) from the equation,
\begin{equation}
D_{\rm S} = D_{\rm Sbh} \left(\frac{H}{H - H_{\rm bh}}\right)^c
\end{equation}
where $D_{\rm S}$ is the basal stem diameter (m), $D_{\rm Sbh}$ stem diameter at breast height, $H_{bh}$ the height at which $D_{\rm Sbh}$ was measured (typically 1.3 or 1.34m). Stem area and diameter were always related assuming the stem was circular. We chose to estimate missing stem diameter rather than stem area because a much better fit was obtained, with a more constant error variance.

The parameter $c$ was further expressed as a function of plant height:
\begin{equation}
c = c_0H^{c_1}.
\end{equation}
The estimated coefficients were $c_0$ = \Sexpr{signif(coef(basalafit)[1],3)}, $c_1$ = \Sexpr{signif(coef(basalafit)[2],3)}, root-mean square error = \Sexpr{signif(summary(basalafit)$sigma,3)}. The relationship was fit using a total of \Sexpr{length(residuals(basalafit))} observations covering the three major PFTs.

@@

From the variance partitioning (Table \ref{table:varparttable}), it is evident that $\afas$ did not vary appreciably between PFTs. Indeed, when all PFTs were plotted together, similar $\af$ was attained at a given $\as$ across the entire plant size range (with the exception of extremely large trees with \textgreater{} 1m$^2$ stem basal area, for which comparatively little data was available) (Fig. 4). We fit a GAM to $\af$ as a function of $\as$ with and without PFT as a covariate, which showed PFT significantly improved model fit (likelihood ratio test, p \textless{} 0.001). Nevertheless, the $R^2$ of the relationship only increased from \Sexpr{round(100*afas$R2a,2)} to \Sexpr{round(100*afas$R2b,2)}\%, indicating that while significant, the PFT had little additional predictive power.

@@

We have confirmed the hypothesis that the amount of leaf area supported per plant is less variable between PFTs than the amount of leaf mass. \citet{poorter2015} hypothesised such a result might arise, because forests would converge to a similar leaf area index (LAI). Although we found remarkably similar leaf area per unit above-ground woody biomass ($\mfms$) between the three PFTs studied (Fig. S2), this does not in itself imply that LAI is equally similar. Indeed, in a global compilation of LAI for undisturbed forest stands, \citet{iio2014} reported a much higher LAI for evergreen gymnosperms compared to evergreen angiosperms, with deciduous angiosperms having intermediate and more variable LAI.  Moreover, we would in fact expect LAI to change with LMA, not because of changes in the amount of leaf area maintained by individual plants, but because LMA influences the shade-tolerance of seedlings \citep{poorter2006, falster2011}, and thereby the LAI of the community \citep{falster2011}. The ultimate reason LMA might effect shade tolerance is because is via its correlation with leaf lifespan \citep{wright2004}. For whole stands, we can therefore expect total stand leaf mass to be correlated with LMA, both because of of greater leaf mass per plant \citep[e.g.]{reich1992}, and the greater density of plants arising from increased leaf lifespan. Previous interpretations of differences in $\mfms$ between gymnosperms and angiosperms have focussed on differences in leaf life-span (LL) \citep{enquistniklas2002, poorter2012}. We argue that there are in fact two effects occurring simultaneously: differences in LMA cause differences in the $\mfms$ of individual plants, while differences in LL (and thereby LMA) cause differences in the LAI of whole stands.
